---
title: "Practical Day One"
author: "Oliver Moses"
format: html
---

`r library(knitr)`

# Question 1

## The rows that contain at least one NA value are as follows:

```{r}
#| echo: true
loop_length <- nrow(airquality)
df_na_rows <- data.frame()
df <- data.frame()
for (i in 1:loop_length) {
  row_vec <- airquality[i, ] 
  if (any(is.na(row_vec))) {
    df_na_rows <- rbind(df_na_rows, row_vec)
  }
  else{
    df <- rbind(df, row_vec)
  }
}
kable(head(df_na_rows, 10))
```

# Question 2

## Ozone and temperature metrics

```{r}
#| echo: true

# Note use DF otherwise it results in some values of NA
avg_ozone <- mean(df$Ozone)
sd_ozone <- sd(df$Ozone)
min_ozone <- min(df$Ozone) 
max_ozone <- max(df$Ozone)

avg_temp <- mean(df$Temp)
sd_temp <- sd(df$Temp)
min_temp <- min(df$Temp) 
max_temp <- max(df$Temp)

final_weather_vec <- data.frame(
  Mean = c(avg_ozone, avg_temp),
  Stand_Dev = c(sd_ozone, sd_temp),
  Min = c(min_ozone, min_temp),
  Max = c(max_ozone, max_temp)
)

kable(final_weather_vec, caption = "The Ozone and Temperature metrcis summarized")
  

```

# Question 3 and Question 4

```{r}
#| echo: true
# Dist = Y, Speed = X
speed_df = cbind(1, cars$speed)
dist_df = matrix(cars$dist)

# Calculating the estimates using the formula ((X^TX)^-1X^TY)
beta_hat <- solve(t(speed_df) %*% speed_df) %*% t(speed_df) %*% dist_df
#print(c("Values using matrix multiplication: ", beta_hat))

# Calculating it using the built in R function
cars_reg = lm(cars$dist~cars$speed)


```

The formula used for the matrix multiplication is: 
\begin{equation}
  \hat{\beta} = (X^TX)^{-1}X^TY 
\end{equation}

The values using matrix multiplication are: `r beta_hat`

The values using the built in LM function are: `r cars_reg$coefficients`

# Question 5 (Jake's additional question):

## The specific values from my LM function are:

```{r}
#| echo: true
mylm <- function(x, y){
  
  # Estimates 
  x_df = cbind(1, x)
  y_df = matrix(y)
  beta_hat <- solve(t(x_df) %*% x_df) %*% t(x_df) %*% y_df
  
  residuals <- y_df - x_df %*% beta_hat
  
  n <- length(y_df)
  p <- ncol(x_df)  # Number of predictors including intercept
  s <- sqrt(sum(residuals^2) / (n - p))

  # Compute standard errors
  var_beta <- s^2 * solve(t(x_df) %*% x_df)  # Variance-covariance matrix
  se_beta <- sqrt(diag(var_beta))  # Standard errors
  
  # T stat
  t_stat <- beta_hat/se_beta
  
  # p_value
  # 2 sided since the test is if beta hat is equal to 0
  deg_of_free = length(y_df) - 2 # number of observations less number of parameters less 1 (n-p-1)
  p_value <- 2 * (1 - pt(abs(t_stat), deg_of_free))
  
  
  
  return(list('Estimate' = beta_hat, 
              'std. Error' = se_beta,
              't value' =  t_stat,
              'Pr(>|t|)' = p_value)
         )
}

a <- mylm(cars$speed, cars$dist)
b <- data.frame(Estimate = a$Estimate,
                Standard_Error = a$`std. Error`,
                T_Value = a$`t value`,
                Pvalue = a$`Pr(>|t|)`
                )
kable(b, caption = "Output from my LM function")
```



